\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for Vietnamese and International Students

### 2023-07-08 phd-cs-us.tex
updates

Several words about stipend. First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost). Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month). Third, like for most source of income in the US, you will have to pay tax on your stipend. Finally, private universities might pay more for stipend.

@@

You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!  You will be very surprised to learn that a school X that you didn't know much about has very strong research in your interested field Y (and vice versa, a school you thought highly off has no one works in the research field you're interested in). This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}. \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.

@@

\section{Acknowledgement}

Many people have contributed to this document.
Craig Yu (GMU) and Hakan Aydin (GMU) provided valuable input in the early version. Other GMU faculty members also have provided feedback and contributions.  Many students including Didier (GMU), Thanh (Melbourne), Dat (Melbourne) have contributed valuable questions and feedback. 
\textbf{Thank you!}

>>>>

Some words stipend. First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost). Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month). Third, like most source of income in the US, you will have to pay tax on your stipend.  Fourth, CS department typically has higher stipend comparing to other study fields.  Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you will have to pay some amount for each class credits).

Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.  In many cases, it is enough to support your spouse and kids (many CS PhD students have their spouse and kids with them). So don't worry too much about stipend.  If you're admitted to a good CS PhD program, you will be fine.

For a full breakdown of how much a graduate student costs, see \S\ref{sec:ra-cost}.

@@

You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!  You will be very surprised to learn that a school that you didn't know much about has very strong research in your interested field (and vice versa, a school you thought highly off has no one works in the research field you're interested in). This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}. \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.

@@

\section{History and Acknowledgement}

\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.  We talked on about why we were not able to attract good Vietnamese international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}) and wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them). I was also  a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.  However, most active participant are students (and some few faculty) in non-CS fields or not in US. Like ChatGPT, their answers to CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion. So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.

I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procastinate :p). I have shared it with many students and faculty members and received many valuable feedback.  I also hope that it will be updated and improved by many people in the future.  I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.  I also hope that this document will be translated to Vietnamese and other languages so that it can reach more students.

\paragraph{Acknowledgement} Many people have contributed to this document.
Craig Yu (GMU) and Hakan Aydin (GMU) provided valuable input in the early version. Other GMU faculty members also have provided feedback and contributions.  Many students including Didier (GMU), Thanh (Melbourne), Dat (Melbourne) have contributed valuable questions and feedback. 
\textbf{Thank you!}
